---
title: "Peer Assessment 1"
author: "Jean Paul Uwimana"
date: "Tuesday, March 10, 2015"
output: html_document
---


```{r}
## Loading and preprocessing the data
setwd("C:/Users/JeanPaul/Documents/ReproducibleResearch/RepData_PeerAssessment1")
if (!(file.exists("./Activity_Monitoring.zip")))
{
    url <- "http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
    download.file(url, destfile = "./Activity_Monitoring.zip")
    unzip("Activity_Monitoring.zip")   
}

activity.data <- read.csv("activity.csv", header = T, sep = ",")
clean.data <- activity.data[complete.cases(activity.data), ] # removing NAs
steps.per.day <- aggregate(clean.data$steps, by = list(clean.data$date), FUN = sum, na.rm = T)
names(steps.per.day) <- c("Date", "TotalSteps") # Renaming columns 
# Printing total steps per day
steps.per.day
png(filename = "totalsteps.png", height = 680, width = 680)
hist(steps.per.day$TotalSteps, main = "Histogram of steps per day", xlab = "Steps per day", col = "red") 
dev.off()
# Median of total steps per day
mean.steps <- aggregate(clean.data$steps, by = list(clean.data$date), FUN = mean)
mean.steps
# Median of total number of steps
median.steps <- aggregate(clean.data$steps, by = list(clean.data$date), FUN = median)
median.steps
```
Time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis):

```{r}
avg.steps <- aggregate(clean.data$steps, by = list(clean.data$interval), FUN = mean)
names(avg.steps) <- c("interval", "steps") # renaming columns to interval & steps, respectively
plot(avg.steps$steps, avg.steps$date, type = "l", main = "5-minute interval vs Average number of steps", xlab = "5-minute interval", ylab = "Average number of steps", col = "blue")
```

The 5-minute interval that contains the maximum number of steps on average across all days:
```{r}
max.steps <- avg.steps[which.max(avg.steps$x), ]
max.steps
```

Total number of missing values
```{r}
missing.values <- activity.data[grepl("NA", activity.data), ]
```

Attempting imputation...
```{r}
#Dataset with missing dataset filled-in
activity.data[is.na(activity.data)] <- avg.steps$steps 
```

Total number of steps taken per day with missing data filled in
```{r}
steps.with.imput <- aggregate(activity.data$steps, by = list(activity.data$date), FUN = sum, na.rm = T)
names(steps.with.imput) <- c("date", "steps")
steps.with.imput # Printing total number of steps with missing data filled in
hist(steps.with.imput$steps, main = "Number of steps per day with missing data filled in", xlab = "Steps per day", col = "red")
```

Mean of total number of steps taken per day with missing data filled in
```{r}
mean.with.imput <- aggregate(activity.data$steps, by = list(activity.data$date), FUN = mean)
mean.with.imput
```

Median of total number of steps taken per day with missing data filled in 
```{r}
median.with.imput <- aggregate(activity.data$steps, by = list(activity.data$date), FUN = median)
median.with.imput
```

Differences in activity between weekdays and weekends
```{r}
# Subsetting for weekdays
weekdays <- activity.data[weekdays(as.Date(activity.data$date)) %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"), ]
weekdays$date <- "weekday" 

# Subsetting for weekends
weekends <- activity.data[weekdays(as.Date(activity.data$date)) %in% c("Saturday", "Sunday"), ]
weekends$date <- "weekend"

# Merging data weekends and weekdays
activity.data <- rbind(weekends, weekdays)
  
# Susbetting by type of day
weekdays <- activity.data[activity.data$date == "weekday", ]
weekends <- activity.data[activity.data$date == "weekend", ]
str(weekdays) # Looking at the dataframe structure
str(weekends) # Looking at the dataframe structure
head(weekdays)
head(weekends)
```
Panel plot of 5-minute interval vs average steps across Weekdays and weekends
```{r, out.width = '680px', out.height = '680px'}
weekdays.mean <- aggregate(weekdays$steps, by = list(weekdays$interval, weekdays$date), FUN = mean) # Averaged weekdays
weekends.mean <- aggregate(weekends$steps, by = list(weekends$interval, weekends$date), FUN = mean) # Averaged weekends
names(weekdays.mean) <- c("interval", "day", "steps")
names(weekends.mean) <- c("interval", "day", "steps")
# Plots below
par(mfrow = c(2, 1))
plot(weekdays.mean$interval, weekdays.mean$steps, type = "l", main = "Weekday", xlab = "5-minute interval", ylab = "Number of steps", col = "blue")
plot(weekends.mean$interval, weekends.mean$steps, type = "l", main = "Weekend", xlab = "5-minute interval", ylab = "Number of steps", col = "blue")
```